---
title: "Data Visualization of Top 5 Cryptocurrencies"
author: "Agustin Espinoza"
date: "2022-11-29"
output: powerpoint_presentation
---

```{r include = FALSE}
library(ggplot2)
library(dplyr)
library(lubridate)
library(scales)
```

## Reserach Question and Background
* How volatile are the top 5 cryptocurrencies on the market?
* Crypto is still fairly new
* Much ongoing research
* This is important to investors who want to make profit and minimize loss

## What is Crypto?
* Digital currency
* Decentralized software system
* Cryptography encrypted
* Mostly Blockchain technology (improving, and better tech emerging)
* Distributed ledger

## Datasets
* Yahoo Finance datasets 
* Use CoinMarketCap API to get real time price data

```{r}
# read csv data
btc_data <- read.csv("BTC-USD-MAX.csv")
eth_data <- read.csv("ETH-USD-MAX.csv")
usdt_data <- read.csv("USDT-USD-MAX.csv")
bnb_data <- read.csv("BNB-USD-MAX.csv")
usdc_data <- read.csv("USDC-USD-MAX.csv")
```

## Hypothesis
* 5 different crypto datasets
* 7 variables per dataset
* 60-100 observations per dataset  
* Hyposthesis: Volatility is inversely proportional to market cap

## Methods
* Error when importing raw CSV files into RStudio
* Solved problem with as.Date() function in R

```{r}
# reformat dates
btc <- btc_data %>%
  mutate(date = ymd(as.Date(Date)))
```

```{r include = FALSE}
# reformat dates (cont.)
eth <- eth_data %>%
  mutate(date = ymd(as.Date(Date)))
usdt <- usdt_data %>%
  mutate(date = ymd(as.Date(Date)))
bnb <- bnb_data %>%
  mutate(date = ymd(as.Date(Date)))
usdc <- usdc_data %>%
  mutate(date = ymd(as.Date(Date)))
```

* Data is already reproducible
* Libraries: ggplot2, dplyr, lubridate, scales

## Bitcoin
* Market Cap = $325 B
* Sept 2014 - Nov 2022

```{r}
# BTC Max Price Plot
ggplot(btc, aes(x=as.Date(Date), y=Close, group=1)) + 
  geom_area(fill="orange", alpha=.4) +
  scale_x_date(date_labels = "%b/%y", breaks = seq.Date(from = as.Date("2014-10-01"), to = as.Date("2022-11-29"), by = 200)) +
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(y= "Bitcoin Price ($)", x = "Date")
```

## Ethereum
* $156 B
* Sept 2015 - Nov 2022

```{r}
# ETH Max Price Plot
ggplot(eth, aes(x=as.Date(Date), y=Close, group=1)) + 
  geom_area(fill="black", alpha=.4) +
  scale_x_date(date_labels = "%b/%y", breaks = seq.Date(from = as.Date("2015-09-01"), to = as.Date("2022-11-30"), by = 200)) +
  scale_y_continuous(labels = label_number(suffix = " K", scale = 1e-3)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(y= "Ethereum Price ($)", x = "Date")
```

## Tether
* $65 B
* Mar 2015 - Nov 2022

```{r}
# USDT Max Price Plot
ggplot(usdt, aes(x=as.Date(Date), y=Close, group=1)) + 
  geom_area(fill="cyan", alpha=.4) +
  scale_x_date(date_labels = "%b/%y", breaks = seq.Date(from = as.Date("2015-03-01"), to = as.Date("2022-11-30"), by = 200)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(y= "Tether Price ($)", x = "Date")
```

## BNB
* $47 B
* Aug 2017 - Nov 2022

```{r}
# BNB Max Price Plot
ggplot(bnb, aes(x=as.Date(Date), y=Close, group=1)) + 
  geom_area(fill="red", alpha=.4) +
  scale_x_date(date_labels = "%b/%y", breaks = seq.Date(from = as.Date("2017-08-01"), to = as.Date("2022-11-30"), by = 200)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(y= "BNB Price ($)", x = "Date")
```

## USD Coin
* $43 B
* Nov 2018 - Nov 2022

```{r}
# USDC Max Price Plot
ggplot(usdc, aes(x=as.Date(Date), y=Close, group=1)) + 
  geom_area(fill="green", alpha=.4) +
  scale_x_date(date_labels = "%b/%y", breaks = seq.Date(from = as.Date("2018-11-01"), to = as.Date("2022-11-30"), by = 200)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(y= "USD Coin Price ($)", x = "Date")
```


## What Did I Learn?
* How to download raw crypto data from reliable sources and APIs
* Visualize data using ggplot2
* Create powerpoint presentations using R Markdown
* More about cryptocurrencies

## Conclusion
* Overall, project was a success
* Did not fully satisfy initial project vision
* However, I still learned a great deal about R and crypto
* Future projects, volatility analysis and crypto trading bot

## References
* Crypto real time prices &amp; latest news Yahoo! Finance. Available at: https://finance.yahoo.com/crypto/ [Accessed December 1, 2022]. 
* Hayes A (2022) Volatility: Meaning in finance and how it works with stocks. Investopedia. Available at: https://www.investopedia.com/terms/v/volatility.asp#:~:text=Volatility%20is%20a%20statistical%20measure%20of%20the%20dispersion%20of%20data,prices%2C%20on%20an%20annualized%20basis. [Accessed December 1, 2022]. 
* Documentation R. Available at: https://www.r-project.org/other-docs.html [Accessed December 1, 2022]. 
